---
title: "Chess_Project_1"
author: "Xhulia Turkaj"
date: "2023-09-24"
output: html_document
---

## Data 607 Project 01 - Chess Tournament Results



In this project, I was given a text file with chess tournament results where the information has some structure. My 
job is to create an R Markdown file that generates a .CSV file (that could for example be imported into a SQL database)
with the following information for all of the players:
Player’s Name, Player’s State, Total Number of Points, Player’s Pre-Rating, and Average Pre Chess Rating of Opponents



### Loading the needed libraries for this project 

```{r}
library(tidyverse)
library(data.table)
library(stringr)
library(dplyr)

```

### Next, loading the txt file from github

```{r}
chess <- read.delim('https://raw.githubusercontent.com/xhuliaturkaj/data607_project_1/main/chess_project_raw.txt', skip = 4, header = FALSE, stringsAsFactors=FALSE)
head(chess)
```

## Formating the current chess data 

chess is now a dataframe with one column V1 and multiple rows

```{r}
#Let's start with removing all the rows that contain hyphens 

chess <- subset(chess, !grepl("^-+$", V1))


```


```{r}
# Next I will combine every two consecutive rows in one single row as each couple of rows contains info per one individual player

str(chess)

chess_combined <- data.frame(V1 = paste(chess$V1[seq(1, nrow(chess), by = 2)], 
                                          chess$V1[seq(2, nrow(chess), by = 2)]))


chess_combined
```
For each observation the information after the arrow -> is unnecessary for my ultimate csv file so I will substitute it with empty string 


```{r}
chess_combined$V1 <- sub("->.*", "", chess_combined$V1)
head(chess_combined$V1)

```

Next I have created the Rounds Columns 
```{r}

column_names <- c("Player_Number", "Player_Name", "Total_Pts", "Round_1", "Round_2", "Round_3", 
                 "Round_4", "Round_5", "Round_6", "Round_7", "State", "Rating")

chess_combined <- chess_combined %>%
separate(V1, into = column_names, sep = "\\|")
                         
```

For the Columns "Round_1", "Round_2", "Round_3", "Round_4", "Round_5", "Round_6" and "Round_7" I only need the numerical values for each observations
```{r}


rounds_columns <- c("Round_1", "Round_2", "Round_3", "Round_4", "Round_5", "Round_6", "Round_7")

#This function will extracte just the numerical values 
extract_numeric <- function(x) {
  as.numeric(str_extract(x, "\\d+"))
}

chess_combined <- chess_combined %>%
  mutate(across(all_of(rounds_columns), extract_numeric))

```

For the Rating column the true rating value is the numerical value after R: so I will extract that value only 

```{r}

chess_combined <- chess_combined %>%
  mutate(Rating = str_extract(str_replace(Rating, "^.*\\/ R: ", ""), "\\d+"))


```

Next I want to change the values of all 7 round columns from the player number to the players rating, this will make the 
Opponent Avg Rating easier to calculate

```{r}
# make sure the rounds columns have numeric values 

chess_combined <- chess_combined %>%
  mutate(across(all_of(rounds_columns), as.numeric))

chess_lookup_table <- data.frame(
  Opponent_Rating = as.numeric(as.character(chess_combined$Rating)), 
  Player_Number = as.numeric(as.character(chess_combined$Player_Number))
)
chess_lookup_table

# Replace the player numbers in round columns with their corresponding ratings

chess_combined <- chess_combined %>%
  rowwise() %>%
  mutate(across(all_of(rounds_columns), 
                 ~ ifelse(is.na(.), NA, 
                          chess_lookup_table$Opponent_Rating[match(., chess_lookup_table$Player_Number)])))


```

Next I will create an extra column that represent the average value of the total Opponents Rating 
```{r}

chess_combined <- chess_combined %>%
  rowwise() %>%
  mutate(Average_Opponent_Rating = mean(c_across(all_of(rounds_columns)), na.rm = TRUE)) %>%
  ungroup()


```

The final step is to subset the df with only our columns of interest and to create a csv file 

```{r}
final_chess_df <- chess_combined %>% select(Player_Name, State,Total_Pts, Rating, Average_Opponent_Rating)%>%
  mutate( Average_Opponent_Rating= round(Average_Opponent_Rating))
```

```{r}
write.csv(final_chess_df, "final_chess_df.csv", row.names = FALSE)

```

